---
title: "Lab 01 - Hello R" 
author: "Sophie Boyd"
date: "1-14-26"
output: 
  html_document: 
    theme: cosmo
    highlight: pygments
---

## Load packages and data

```{r load-packages, message=FALSE}
library(tidyverse) 
library(datasauRus)
data("datasaurus_dozen")
```

## Exercises

### Exercise 1

The file has 1846 rows and 3 columns. The variables are dataset (which dataset the data come from), x (x-values), and y (y-values).

### Exercise 2

#### Visualizing the dino data:

```{r plot-dino, fig.height=3, fig.width=6}
dino_data <- datasaurus_dozen %>%
  filter(dataset == "dino")

ggplot(data = dino_data, mapping = aes(x = x, y = y)) +
  geom_point()
```

#### Correlation between `x` and `y` in the dino dataset:

```{r cor-dino}
dino_data %>%
  summarize(r = cor(x, y))
```

On its own, the correlation between x and y might be interpreted as a weak, negative relationship between the two variables. However, because we plotted the data, we can see that x and y do not have a linear relationship. Rather, the data points make the shape of a dinosaur! This exercise highlights the importance of plotting/visualizing your data before conducting analyses and drawing conclusions. 

### Exercise 3

#### Visualizing the star data:

```{r plot-star, fig.height=3, fig.width=6}
star_data <- datasaurus_dozen %>%
  filter(dataset == "star")

ggplot(data = star_data, mapping = aes(x = x, y = y)) +
  geom_point()
```

As with the dino data, x and y do not have a linear relationship. Here, the plotted data takes the shape of a star.



#### Correlation between `x` and `y` in the star dataset:

```{r cor-star}
star_data %>%
  summarize(r = cor(x, y))
```
Again, on its own this correlation might be interpreted as a weak negative relationship between x and y, but it should not be interpreted this way because the plot does not show a linear relationship. The correlation coefficient is very close to the r value that we obtained from the dino data (both are about -.06), demonstrating that very different datasets may produce similar summary statistics (as the creator intended!).

### Exercise 4

#### Visualizing the circle data:

```{r plot-circle, fig.height=3, fig.width=6}
circle_data <- datasaurus_dozen %>%
  filter(dataset == "circle")

ggplot(data = circle_data, mapping = aes(x = x, y = y)) +
  geom_point()
```

#### Correlation between `x` and `y` in the circle dataset:

```{r cor-circle}
circle_data %>%
  summarize(r = cor(x, y))
```
The correlation of -.068 between x and y in the circle dataset is very close to the correlation coefficients produced by both the dino and star datasets (both about -.06), once again demonstrating that very different datasets can produce similar summary statistics. 

### Exercise 5

#### Visualizing all of the datasets together:

```{r plot-datasaurus, fig.height=6, fig.width=6}
ggplot(datasaurus_dozen, aes(x = x, y = y, color = dataset)) +
  geom_point() +
  facet_wrap(~dataset, ncol = 3) +
  theme(legend.position = "none")
```

#### Correlations between `x` and `y` in all of the datasets:

```{r cor-datasaurus}
datasaurus_dozen %>%
  group_by(dataset) %>%
  summarize(r = cor(x, y)) %>%
  print(13)
```
Across all 13 datasets contained in datasaurus, we see that plotting x and y produces drastically different patterns, but the correlations between x and y are nearly identical across the datasets. The main takeaway from lab 1: always plot your data first! Data visualization allows us to determine the appropriate analyses to conduct and to draw responsible conclusions from our analyses.


